---
title: "RA Project on Body Fat"
author: "Charmi, Rishi, Harsh"
date: "2023-04-07"
output: html_document
---

# IMPORTING LIBRARIES

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(ggpubr)
library(ggplot2)
library(broom)
library(tidyverse)
library(corrplot)
library(lmtest)
library(PerformanceAnalytics)
library(car)
library(DescTools)
```

# INTRODUCTORY EXPLORATION

## IMPORTING DATASET

```{r}
body.fat <- read.csv("D:/NMIMS/Sem.4/Projects/Body Fat.csv")
head(body.fat)
```

> The dataset contains 252 observations and 15 variables.

Quick glance of data, our target variable is Density of fat in a person and other variables may be parameter. The data

## OBJECTIVE

> The objective of this study is to understand the relationship between the target variable Density and its relationship with the body fat percentage and different sizes of body parts using Multiple Linear Regression Model(MLR).

## VARIABLES AND THEIR TYPE

```{r}
sapply(body.fat, typeof)
```

**Continuous Variable**: All the variables are continuous.

## DESCRIPTION OF EACH VARIABLE

1.  Density(ordinal) - Density of the fat in the body
2.  BodyFat(ordinal) - Percentage of body fat
3.  Age(ordinal) - Age(year)
4.  Weight(ordinal) - Weight(lbs)
5.  Height(ordinal) - Height(inches)
6.  Neck(ordinal) - Neck size(cm)
7.  Chest(ordinal) - Chest size(cm)
8.  Abdomen(ordinal) - Abdomen size(cm)
9.  Hip(ordinal) - Hip size(cm)
10. Thigh(ordinal) - Thigh size(cm)
11. Knee(ordinal) - Knee size(cm)
12. Ankle(ordinal) - Ankle size(cm)
13. Biceps(ordinal) - Biceps size(cm)
14. Forearm(ordinal) - Forearm size(cm)
15. Wrist(ordinal) - Wrist size(cm)

## A QUICK SUMMARY

```{r}
summary(body.fat)
```

```{r}
str(body.fat)
```

## OBSERVATION

**Density:** The data range from 0.995 to 1.109, with an average value of 1.056, indicating that most of the measurements cluster around this value. The median value of 1.055 indicates that half of the measurements fall below this value, while the other half falls above it.

**Body Fat:** The data range from 0.00% to 47.50%, with an average value of 19.15% and a median value of 19.20%, indicating that the distribution is relatively symmetrical.

**Age:** The data range from 22 to 81 years, with an average age of 44.88 years and a median age of 43 years. This indicates that the majority of participants fall within this age range.

**Weight:** The data range from 118.5 lbs to 363.1 lbs, with an average weight of 178.9 lbs and a median weight of 176.5 lbs, indicating that the distribution is slightly skewed to the right.

**Height:** The data range from 29.50 inches to 77.75 inches, with an average height of 70.15 inches and a median height of 70.00 inches, indicating that the distribution is relatively symmetrical.

**Neck:** The data range from 31.10 cm to 51.20 cm, with an average value of 37.99 cm and a median value of 38.00 cm, indicating that the distribution is relatively symmetrical.

**Chest:** The data range from 79.30 cm to 136.20 cm, with an average value of 100.82 cm and a median value of 99.65 cm, indicating that the distribution is slightly skewed to the right.

**Abdomen:** The data range from 69.40 cm to 148.10 cm, with an average value of 92.56 cm and a median value of 90.95 cm, indicating that the distribution is slightly skewed to the right.

**Hip:** The data range from 85.0 cm to 147.7 cm, with an average value of 99.9 cm and a median value of 99.3 cm, indicating that the distribution is relatively symmetrical.

**Thigh:** The data range from 47.20 cm to 87.30 cm, with an average value of 59.41 cm and a median value of 59.00 cm, indicating that the distribution is relatively symmetrical.

**Knee:** The data range from 33.00 cm to 49.10 cm, with an average value of 38.59 cm and a median value of 38.50 cm, indicating that the distribution is relatively symmetrical.

**Ankle:** The data range from 19.1 cm to 33.9 cm, with an average value of 23.1 cm and a median value of 22.8 cm, indicating that the distribution is relatively symmetrical.

**Biceps:** The data range from 24.80 cm to 45.00 cm, with an average value of 32.27 cm and a median value of 32.05 cm, indicating that the distribution is relatively symmetrical.

**Forearm:** The data range from 21.00 cm to 34.90 cm, with an average value of 28.66 cm and a median value of 28.70 cm, indicating that the distribution is relatively symmetrical.

**Wrist:** The data range from 15.80 cm to 21.40 cm, with an average value of 18.23 cm and a median value of 18.30 cm, indicating that the distribution is relatively symmetrical.

## MISSING VALUE TREATMENT

Finding Null Values

```{r}
colSums(is.na(body.fat))
colSums(is.na(body.fat))[colSums(is.na(body.fat)) > 0]
rowSums(is.na(body.fat))
```

Their are no null values in the dataset.

## OUTLIER DETECTION AND REMOVAL

Outliers are extreme values that are outside the range of what is expected and unlike the other data.<br> Detecting and removing outliers can improve the accuracy of the model.

```{r}
boxplot(body.fat$BodyFat)
boxplot(body.fat$Age)
boxplot(body.fat$Weight)
boxplot(body.fat$Height)
boxplot(body.fat$Neck)
boxplot(body.fat$Chest)
boxplot(body.fat$Abdomen)
boxplot(body.fat$Hip)
boxplot(body.fat$Thigh)
boxplot(body.fat$Knee)
boxplot(body.fat$Ankle)
boxplot(body.fat$Biceps)
boxplot(body.fat$Forearm)
boxplot(body.fat$Wrist)
```

Observing the boxplots, there are significant amount of outliers in each column.<br> Thus they can't be discarded and have to be dealt with using Winsorizing.

Winsorizing is a method of outlier treatment wherein the values beyond the extreme value limits are imputed by those limits.<br> This extreme value is detected as any point beyond a certain percentile.

```{r}
body.fat[2:15] <- lapply(body.fat[2:15], Winsorize)
```

# CHECKING THE ASSUMPTIONS

## ASSUMPTION.1: LINEARITY

> Linear relationship between dependent and independent variable.

We can check this assumption by plotting each independent variable against dependent variable in a scatter plot.

```{r}
par(mfrow = c(2,2))
plot(body.fat$Density ~ body.fat$BodyFat)
plot(body.fat$Density ~ body.fat$Age)
plot(body.fat$Density ~ body.fat$Weight)
plot(body.fat$Density ~ body.fat$Height)
plot(body.fat$Density ~ body.fat$Neck)
plot(body.fat$Density ~ body.fat$Chest)
plot(body.fat$Density ~ body.fat$Abdomen)
plot(body.fat$Density ~ body.fat$Hip)
plot(body.fat$Density ~ body.fat$Thigh)
plot(body.fat$Density ~ body.fat$Knee)
plot(body.fat$Density ~ body.fat$Ankle)
plot(body.fat$Density ~ body.fat$Biceps)
plot(body.fat$Density ~ body.fat$Forearm)
plot(body.fat$Density ~ body.fat$Wrist)
```

The relationship looks roughly linear, so we can proceed with the linear
model.

## ASSUMPTION.2: INDEPENDENT VARIABLES (MULTI-COLINEARITY)

> The independent variables are not highly co-related.

We can check this assumption by calculating the correlation among the independent variable.

```{r}
correlation <- cor(body.fat[,2:15])
correlation > 0.7
```

So variable Weight, Chest, Abdomen, Hip, Thigh seems to have the higher correlations with all other variables.

```{r}
corrplot(correlation)
```

```{r}
chart.Correlation(correlation)
```

```{r}
heatmap(correlation)
```

Variance Inflation Factor(VIF) measures how much the behavior of an independent variable is influenced by its correlation with the other independent variables.<br> It allows a quick measure of how much a variable is contributing to the standard error in the regression.<br> Generally VIF greater than 5 suggests multi-colinearity.

```{r}
model.lm <- lm( body.fat$Density ~ .,  data = body.fat )
VIF(model.lm)[VIF(model.lm) > 5]
```

```{r}
model.lm <- lm( Density ~ BodyFat + Age + Height + Neck + Chest + Abdomen + Hip + Thigh + Knee + Ankle + Biceps + Forearm + Wrist, data = body.fat )
VIF(model.lm)[VIF(model.lm) > 5]
```

```{r}
model.lm <- lm( Density ~ BodyFat + Age + Height + Neck + Chest + Hip + Thigh + Knee + Ankle + Biceps + Forearm + Wrist, data = body.fat )
VIF(model.lm)[VIF(model.lm) > 5]
```

```{r}
model.lm <- lm( Density ~ BodyFat + Age + Height + Neck + Chest + Thigh + Knee + Ankle + Biceps + Forearm + Wrist, data = body.fat )
VIF(model.lm)[VIF(model.lm) > 5]
```

```{r}
model.lm <- lm( Density ~ BodyFat + Age + Height + Neck + Chest + Knee + Ankle + Biceps + Forearm + Wrist, data = body.fat )
summary(model.lm)
```

## ASSUMPTION.3: VARIANCE (HOMOSCEDASTICITY)

> The variance of residuals is constant.

The variance of residuals is approximately constant and normal.<br> Commented using "Residuals vs Fitted" and "normal QQ plot"

```{r}
par(mfrow = c(2,2))
plot(model.lm)
```

Homoscedasticity is a statistical assumption that the variance of the errors or residuals in a regression model is constant across all levels of the independent variables.<br> In other words, it means that the spread of the residuals is the same across the range of predicted values. Homoscedasticity is important because it ensures that the standard errors of the regression coefficients are accurate, which is necessary for making valid inferences about the population parameters.<br> Violations of homoscedasticity can lead to biased and inefficient estimates of the regression coefficients and can affect the validity of statistical tests.

Breusch-Pagan test is that the residuals have constant variance, while the alternative hypothesis is that the residuals have non-constant variance.

```{r}
bptest(model.lm)
```

The test statistic (BP) is 13.577 and the degrees of freedom (df) is 10. The p-value associated with the test is 0.1932, which is greater than the significance level of 0.05.<br>
Since the p-value is greater than the significance level, we fail to reject the null hypothesis. Therefore, we can conclude that there is no evidence of heteroscedasticity in the regression model at the 5% significance level.

## ASSUMPTION.4: INDEPENDENCE OF OBSERVATIONS (AUTOCORRELATION)

> There should be independence of observations.

The Durbin-Watson test is a statistical test used to detect the presence of autocorrelation in the residuals of a regression model. The test statistic ranges positive from 0 to 4, with a value of 2 indicating no autocorrelation.<br> A value of d \< 2 indicates serial correlation, while a value of d \> 2 indicates negative serial correlation.<br> The test iscommonly used in time series analysis to check for autocorrelation in the residuals.

```{r}
dwtest(model.lm)
```

The DW statistic ranges from zero to four, with a value of 2.0 indicating zero autocorrelation.<br>
Values below 2.0 mean there is positive autocorrelation and above 2.0 indicates negative autocorrelation.<br>
The DW value is closer to 2 and the p-value is greater than level of significance i.e. 0.05, we fail to reject null hypothesis.

## ASSUMPTION.5: MULTIVARIATE NORMALITY

> Residual a re normally distributed.

*Using Shapiro Test*

The Shapiro-Wilk test is a statistical test used to determine whether a data set is normally distributed or not.<br> The test uses the following hypotheses,<br> H0 (null hypothesis): The data is normally distributed.<br> HA (alternative hypothesis): The data is not normally distributed.<br> The test is widely used in various fields, including finance, biology, psychology, and social sciences, to name a few.<br> If the p-value of the Shapiro-Wilk test is less than the significance level (usually 0.05), then we reject the null hypothesis and conclude that the data is not normally distributed.

```{r}
shapiro.test(model.lm$residuals)
```

*Using Histogram*

```{r}
par(mfrow = c(1,1))
h <- hist(body.fat$Age)
xfit <- seq(min(body.fat$Age), max(body.fat$Age), length = 40) 
yfit <- dnorm(xfit, mean = mean(body.fat$Age), sd = sd(body.fat$Age)) 
yfit <- yfit * diff(h$mids) * length(body.fat$Age) 
lines(xfit, yfit, col = "darkblue", lwd = 2)
```

It appears that the data is roughly normally distributed, as the fitted curve appears to match the shape of the histogram fairly well.

# MODEL

```{r}
model.lm <- lm( Density ~ BodyFat + Age + Height + Neck + Chest + Knee + Ankle + Biceps + Forearm + Wrist, data = body.fat )
summary(model.lm)
```

# POTTING THE REGRESSION LINE

```{r}
ggplot(body.fat, aes(x =BodyFat, y = Density)) +
  geom_point() +
  xlab("BodyFat") +
  ylab("Density") +
  ggtitle("Scatter plot of Density vs BodyFat")+
  geom_smooth(method = "lm", se = FALSE)
```

# INTERPRETATION

> Reporting the results of Multiple Linear Regression

In this study of 252 observations, we found the coefficient for "BodyFat" is estimated to be -0.002377, which means that for every one-unit increase in "BodyFat", the predicted value of "Density" decreases by 0.002377 units.<br> This coefficient is highly statistically significant (p \< 0.001), indicating that "BodyFat" is a very important predictor of "Density" in this model.<br> The intercept of the regression line is estimated to be 1.094, meaning that when all predictor variables are equal to zero, the predicted value of "Density" is 1.094.
